---
title: "Monthly Weather Generator"
author: "Jeffrey D Walker, PhD"
date: "August 21, 2014"
output:
  html_document:
    toc: yes
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Monthly Weather Generator}
-->

# Load Libraries

```{r libraries, warning=FALSE, message=FALSE}
library(forecast)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(ggplot2)
theme_set(theme_bw())
library(weatherGen)
data(maurer)

# set RNG seed for reproducibility
set.seed(2345)
```

# Load Monthly Climate Data

This analysis will use the [gridded climate dataset by Maurer et al, 2002](http://www.engr.scu.edu/~emaurer/gridded_obs/index_gridded_obs.html), which is provided in the `maurer` dataset in this package. This dataset was created using the makefile and R scripts in the `/scripts/maurer_mon` directory of the package.

Here is a map of the climate data grid points.

```{r map east, warning=FALSE}
df.grid <- maurer$grid
library(ggmap)
map <- get_map(location=c(lon=mean(range(df.grid$LON)), 
                          lat=mean(range(df.grid$LAT))),
               zoom=6, maptype="satellite", color='bw')
ggmap(map, darken=c(0.25, "white"), extent="device") +
  geom_point(aes(x=LON, y=LAT), data=df.grid, color='red', size=1)
```

Zoomed into new england. The green points will be used to compute a regional average.

```{r map new england, warning=FALSE}
ggmap(map, darken=c(0.25, "white"), extent="device") +
  geom_point(aes(x=LON, y=LAT, color=SELECT), 
             data=mutate(df.grid, SELECT=(LAT<=43)), size=1) +
  geom_hline(yint=43, color='green', linetype=2) +
  scale_color_manual(values=c("TRUE"="green", "FALSE"="red"), guide=FALSE)
```

## Monthly Climate Dataset

Using the selected grid points, retrieve the monthly timeseries from the database. Also, compute the variable `TAVG` as the arithmetic mean of `TMIN` and `TMAX`.

```{r climate load}
clim.mon <- maurer$data %>%
  filter(LAT<=43) %>%
  collect %>%
  select(-WIND) %>%
  mutate(TAVG=(TMIN+TMAX)/2) %>%
  gather(VAR, VALUE, PRCP, TMIN, TMAX, TAVG)
summary(clim.mon)
```

Compute annual timeseries of each climate variable and each location. Note that precipitation is simply the annual sum, while the other three variables are computed as weighted means based on the number of days in each month to get a correct annual mean.

```{r climate yr}
clim.yr <- spread(clim.mon, VAR, VALUE) %>%
  mutate(MONTHYEAR=ymd(paste(YEAR,MONTH,1,sep='-')),
         N_DAY=days_in_month(MONTHYEAR)) %>%
  group_by(LAT, LON, YEAR) %>%
  summarise(PRCP=sum(PRCP),
            TMAX=sum(TMAX*N_DAY)/sum(N_DAY),
            TMIN=sum(TMIN*N_DAY)/sum(N_DAY),
            TAVG=sum(TAVG*N_DAY)/sum(N_DAY)) %>%
  gather(VAR, VALUE, PRCP:TAVG)
summary(clim.yr)
```


## Regional Average

Compute the regional average monthly value of each climate variable.

```{r climate reg mon}
clim.reg.mon <- clim.mon %>%
  group_by(YEAR, MONTH, DATE, VAR) %>%
  summarise(MEAN=mean(VALUE),
            SD=sd(VALUE))
summary(clim.reg.mon)
```

This figure shows the regional average monthly climate timeseries. The shaded region shows mean +/- 1 standard deviation.

```{r plot clim reg mon}
clim.reg.mon %>%
  ggplot(aes(DATE, MEAN)) +
  geom_ribbon(aes(ymin=MEAN-SD, ymax=MEAN+SD), fill='grey50') +
  geom_line(color='blue') +
  facet_wrap(~VAR, ncol=1, scales='free_y') +
  labs(x="Month/Year", y="Mean +/- StDev")
```

Compute the regional average annual value of each climate variable.

```{r climate reg yr}
clim.reg.yr <- clim.yr %>%
  group_by(VAR, YEAR) %>%
  summarise(MEAN=mean(VALUE),
            SD=sd(VALUE))
summary(clim.reg.yr)
```

This figure shows the regional-average annual climate timeseries. The shaded region shows the mean +/- 1 standard deviation.

```{r plot clim reg yr}
ggplot(clim.reg.yr, aes(YEAR, MEAN)) +
  geom_ribbon(aes(ymin=MEAN-SD, ymax=MEAN+SD), fill='grey80') +
  geom_line(color='blue') +
  facet_wrap(~VAR, ncol=1, scales='free_y') +
  labs(x="Year", y="Mean +/- StDev")
```

Convert the regional average datasets wide format for use in the weather generator.

```{r clim reg wide}
clim.reg.mon <- select(clim.reg.mon, -SD) %>%
  spread(VAR, MEAN)
head(clim.reg.mon)

clim.reg.yr <- select(clim.reg.yr, -SD) %>%
  spread(VAR, MEAN)
head(clim.reg.yr)
```

# Wavelet Analysis

Perform wavelet analysis on regional average annual precipitation timeseries.

```{r plot wave, message=FALSE, warning=FALSE}
wt.reg <- wavelet_analysis(clim.reg.yr$PRCP, sig.level=0.90, noise.type='white')

par(mfrow=c(1,2))
plot(wt.reg, plot.cb=TRUE, plot.phase=FALSE)
plot(wt.reg$gws, wt.reg$period, type="b",
     xlab="Global Wavelet Spectrum", ylab="",
     log="y", ylim=rev(range(wt.reg$period)), 
     xlim=range(c(wt.reg$gws, wt.reg$gws.sig$signif)))
lines(wt.reg$gws.sig$signif, wt.reg$period, lty=2, col='red')
```

Although there are some significant wavelet periods, we'll use a simple arima model of the regional annual precipitation instead of arima models of the wavelet components.

```{r arima model}
models <- list(PRCP=auto.arima(clim.reg.yr$PRCP,
                               max.p=2,max.q=2,max.P=0,max.Q=0,
                               stationary=TRUE))
models[["PRCP"]]
```

This figure shows a 20-year forecast of the arima model reflecting no low-frequency oscillations (since we are not using the wavelet decomposition).

```{r plot arima forecast}
par(mfrow=c(1,1))
plot(forecast(models[['PRCP']], h=20), main="ARIMA Forecast of Regional Annual Precip",
     xlab="Timestep (yr)", ylab='Annual Precip (mm/yr)')
```

Generate 50 simulations of annual precipitation using the AR1 model of the regional annual precipitation.

```{r mc ar, warning=FALSE}
sim.prcp <- weatherGen::mc_arimas(models=models, n=nrow(clim.reg.yr), n.iter=50)
str(sim.prcp)
```

Compare the GWS power and the mean, standard deviation, and skewness statistics across these simulations (red point is the overall mean, boxplots are the distribution across simulations) to regional annual precipitation timeseries (blue point).  

```{r plot gws stats, fig.height=8}
par(mfrow=c(2,2))
x1 <- 1
x2 <- min(length(wt.reg$period), nrow(sim.prcp$gws))
ymin <- min(wt.reg$period[x1:x2], wt.reg$gws[x1:x2],
            sim.prcp$gws.stat[x1:x2, 'MEAN'],
            sim.prcp$gws.stat[x1:x2, 'Q025'])
ymax <- max(wt.reg$period[x1:x2], wt.reg$gws[x1:x2],
            sim.prcp$gws.stat[x1:x2, 'MEAN'],
            sim.prcp$gws.stat[x1:x2, 'Q975'])
plot(wt.reg$period[x1:x2], wt.reg$gws[x1:x2],
     type="n", ylim=c(ymin,ymax), xlab="Period (years)", ylab="",
     main="", log="y")
mtext(side=2, expression(paste("Power (",mm^2,")")), line=2.5)
polygon(c(wt.reg$period[x1:x2],
          rev(wt.reg$period[x1:x2])),
        c(sim.prcp$gws.stat[x1:x2, 'Q025'],
          rev(sim.prcp$gws.stat[x1:x2, 'Q975'])),
        col="grey")
lines(wt.reg$period[x1:x2], wt.reg$gws[x1:x2])
lines(wt.reg$period[x1:x2], sim.prcp$gws.stat[x1:x2, 'MEAN'], lty=2, col="blue")
lines(wt.reg$period[x1:x2], wt.reg$gws.sig$signif[x1:x2], col="red", lty=3)

boxplot(colMeans(sim.prcp$x), main="Mean")
points(mean(clim.reg.yr$PRCP),col="red",pch=19)
points(mean(colMeans(sim.prcp$x)), col="blue", pch=19)

boxplot(apply(sim.prcp$x, 2, sd), main="Standard Deviation")
points(sd(clim.reg.yr$PRCP), col="red", pch=19)
points(mean(apply(sim.prcp$x, 2, sd)), col="blue", pch=19)

boxplot(apply(sim.prcp$x, 2, skewness), main="Skew")
points(skewness(clim.reg.yr$PRCP), col="red", pch=19)
points(mean(apply(sim.prcp$x, 2, skewness)),col="blue",pch=19)
```

This figure shows the simulated and observed regional-average annual precipitation. The simulations were generated using the AR1 model described above.

```{r plot ar sims}
as.data.frame(sim.prcp$x) %>%
  mutate(TIMESTEP=row_number()) %>% 
  gather(TRIAL, VALUE, -TIMESTEP) %>% 
  ggplot(aes(TIMESTEP, VALUE)) + 
  geom_line(aes(group=TRIAL), alpha=0.3) + 
  stat_summary(fun.y=mean, geom='line', color='red', size=1) +
  geom_line(aes(x=TIMESTEP, y=PRCP), data=mutate(clim.reg.yr, TIMESTEP=row_number()), color='blue', size=1) +
  labs(x="Year", y="Annual Precipitation (mm/yr)", 
       title="WARM Annual Precipitation Simulations\nRed Line=Mean of Simulations, Blue Line=Observd Regional Mean")
```

# Monthly Weather Generator

## Select Location

First, select a random location from the complete climate dataset.

```{r select loc}
clim.locs <- select(clim.yr, LAT, LON) %>%
  unique()
loc <- clim.locs[sample(nrow(clim.locs), size=1),] %>% as.numeric
names(loc) <- c('LAT', 'LON')
loc
```

Now extract the monthly and annual climate datasets for this location.

```{r loc datasets}
clim.loc.yr <- filter(clim.yr, LAT==loc[['LAT']], LON==loc[['LON']]) %>%
  spread(VAR, VALUE) %>%
  arrange(YEAR)
clim.loc.mon <- filter(clim.mon, LAT==loc[['LAT']], LON==loc[['LON']]) %>%
  spread(VAR, VALUE) %>%
  arrange(YEAR, MONTH)
```



This figure shows the original climate timeseries for the regional-average dataset and the selected location.

```{r plot clim loc reg, fig.height=8}
clim.loc.mon %>%
  select(-LAT, -LON) %>%
  mutate(SOURCE="Location") %>%
  rbind(clim.reg.mon %>% mutate(SOURCE="Region")) %>%
  mutate(SOURCE=factor(SOURCE)) %>%
  gather(VAR, VALUE, PRCP:TAVG) %>%
  mutate(DATE=ymd(paste(YEAR, MONTH, 1, sep='-'))) %>%
  ggplot(aes(DATE, VALUE, color=SOURCE)) +
  geom_line() +
  labs(x="Month/Year", y="") +
  scale_color_manual('', values=c('steelblue', 'orangered')) +
  facet_wrap(~VAR, ncol=1, scales='free_y')
```


## Nearest Neighbors

Compute the euclidian distance between a single simulated annual precipitation value and each observed year. The simulated annual precipitation is for the first year in the first simulation trial.

```{r distances}
j <- 1 # year number
samp <- 1 # simulation trial number

# compute distances between simulated WARM timeseries and observed regional precip
distances <- cbind(clim.reg.yr, DISTANCE=sqrt((sim.prcp$x[j,samp] - clim.reg.yr$PRCP)^2))

ggplot(distances, aes(YEAR, PRCP)) + 
  geom_line() +
  geom_point(aes(color=DISTANCE), size=3) +
  geom_hline(yint=sim.prcp$x[j,samp], linetype=2, color='red') +
  geom_text(x=max(distances$YEAR), y=sim.prcp$x[j,samp], label="Sim", hjust=1, vjust=-0.5) +
  scale_color_gradient(low='green', high='red') +
  labs(x="Year", y="Annual Precipitation (mm/yr)", 
       title="Regional and Current Simulation Annual Precipitation")
```

Select the eight years from the observed regional annual simulation that are the most similar to the first simulation trial and year.

```{r plot top distances}
distances <- mutate(distances,
                    IN_TOP_8=rank(DISTANCE) %in% seq(1,8))
distances %>%
  ggplot(aes(YEAR, PRCP)) + 
  geom_line() +
  geom_point(aes(color=DISTANCE), size=3) +
  geom_hline(yint=sim.prcp$x[j,samp], linetype=2, color='red') +
  geom_point(mapping=aes(alpha=IN_TOP_8), color='black', shape=1, size=4) +
  geom_text(aes(x=YEAR, y=PRCP, label=YEAR), data=filter(distances, IN_TOP_8), vjust=-1) +
  geom_text(x=max(distances$YEAR), y=sim.prcp$x[j,samp], label="Sim", hjust=1, vjust=-0.5) +
  scale_alpha_manual(values=c("TRUE"=1, "FALSE"=0), guide=FALSE) +
  scale_color_gradient(low='green', high='red') +
  labs(x="Year", y="Annual Precipitation (mm)", title="Regional Annual Precipitation with 8 Nearest Neighbors Selected")
```

Extract the observed regional precipitation for the candidate years and compute the sampling weights as inverse distance-squared weighted.

```{r top distances}
distances.top <- filter(distances, IN_TOP_8) %>%
  mutate(WEIGHT=(1/DISTANCE)^2,
         WEIGHT=WEIGHT/sum(WEIGHT))
distances.top %>%
  ggplot(aes(PRCP, WEIGHT)) +
  geom_point() +
  geom_vline(xint=sim.prcp$x[j,samp], linetype=2, color='red') +
  geom_text(aes(label=YEAR), hjust=-0.1) +
  geom_text(x=sim.prcp$x[j,samp], y=0.4, label="Sim Precip", hjust=-0.1) +
  labs(x="Annual Precipitation (mm)", y="Sample Weight")
```

Now we can randomly sample from the 8 nearest neighbors using the sampling weights based on the inverse distance squared. We'll do this 1000 times and compare the frequency of the samples to the sample weights.

```{r sample years}
sampled.years <- sample(distances.top$YEAR,
                        size=1000,
                        prob=distances.top$WEIGHT,
                        replace=TRUE)
sampled.years.tbl <- prop.table(table(sampled.years))
merge(distances.top, 
      data.frame(YEAR=names(sampled.years.tbl),
                 FREQ=as.vector(sampled.years.tbl)),
      all.x=TRUE) %>%
  select(YEAR, WEIGHT, FREQ) %>%
  ggplot(aes(WEIGHT, FREQ)) +
  geom_point(size=3) +
  geom_abline(linetype=2) +
  labs(x="Sample Weight", y="Sampled Frequency")
```

Now if we just sample for one year.

```{r sample year}
sampled.year <- sample(distances.top$YEAR, size=1, prob=distances.top$WEIGHT, replace=TRUE)
sampled.year
```

Then we can extract the observed monthly climate data for the sampled year from the observed dataset of the selected location.

```{r sampled monthly climate}
filter(clim.loc.mon, YEAR==sampled.year)
```

## Generate Complete Timeseries

Given the selected location, repeat the nearest neighbor process for each of the `num_sim` simulations and `num_year` years, and store the result in a 3-dimensional array.

```{r gen set up}
num_sim <- ncol(sim.prcp$x)
num_year <- nrow(sim.prcp$x) # number of composited years
num_months <- num_year*12

climate_variables <- select(clim.reg.yr, -YEAR) %>%
  names()

sampled.year <- array(NA, c(num_year, num_sim))
gen.loc.mon.arr <- array(NA, c(num_months, length(climate_variables), num_sim))
```

```{r gen loop}
# loop through simulations (trials)
for (samp in 1:num_sim) {
    # loop through years
    for (j in 1:num_year) {
        # compute distances between simulated WARM timeseries and observed regional precip
      distances <- mutate(clim.reg.yr, 
                          DISTANCE=sqrt((sim.prcp$x[j, samp] - clim.reg.yr$PRCP)^2),
                          IN_TOP_8=rank(DISTANCE) %in% seq(1,8))
      distances.top <- filter(distances, IN_TOP_8) %>%
                       mutate(WEIGHT=(1/DISTANCE)^2,
                              WEIGHT=WEIGHT/sum(WEIGHT))
      sampled.year[j,samp] <- sample(distances.top$YEAR,
                                     size=1,
                                     prob=distances.top$WEIGHT,
                                     replace=FALSE)
      row.idx <- ((1+12*(j-1)):(12+12*(j-1)))
      gen.loc.mon.arr[row.idx, , samp] <- data.matrix(filter(clim.loc.mon, 
                                                             YEAR==sampled.year[j, samp]) %>%
                                                      select(PRCP, TMAX, TMIN, TAVG))
    }
}
```

Now convert the 3-dim array to a list of dataframes with each element in the list containing the simulated data for a single variable.

```{r gen loc mon}
gen.loc.mon <- sapply(climate_variables, function(v) {
  i.var <- which(climate_variables==v)
  df <- as.data.frame(gen.loc.mon.arr[,i.var,]) %>%
    mutate(TIMESTEP=row_number(),
           SIM_YEAR=rep(seq(1, num_year), each=12),
           SIM_MONTH=rep(seq(1, 12), times=num_year),
           VAR=v) %>%
    gather(TRIAL, VALUE, -TIMESTEP, -SIM_YEAR, -SIM_MONTH, -VAR) %>%
    mutate(TRIAL=as.numeric(TRIAL))
  return(list(df))
})
str(gen.loc.mon)
```

Aggregate the generated timeseries to compute annual values or each variable.

```{r gen loc yr}
gen.loc.yr <- lapply(gen.loc.mon, function(df) {
  if ('PRCP' %in% unique(df$VAR)) {
    df.yr <- group_by(df, VAR, TRIAL, SIM_YEAR) %>%
      summarise(VALUE=sum(VALUE))
  } else {
    df.yr <- group_by(df, VAR, TRIAL, SIM_YEAR) %>%
      summarise(VALUE=mean(VALUE))
  }
  df.yr <- df.yr %>%
    ungroup() %>%
    as.data.frame()
  return(df.yr)
})
str(gen.loc.yr)
```

This figure shows all the weather generator simulations compared to the observed climate dataset at the selected location.

```{r plot gen loc yr}
do.call(rbind, gen.loc.yr) %>%
  mutate(VAR=factor(VAR)) %>%
  ggplot() +
    geom_line(aes(SIM_YEAR, VALUE, group=TRIAL), alpha=0.5) +
    geom_line(aes(SIM_YEAR, VALUE), 
              data=mutate(clim.reg.yr, SIM_YEAR=row_number()) %>%
                gather(VAR, VALUE, PRCP:TAVG), 
              color='red') +
    facet_wrap(~VAR, scales='free_y') +
    labs(x="Simulation Year", y="Value", title="All Generated Datasets with Observed Location Dataset ")
```

This figure shows only the first weather generator simulation compared to the observed climate dataset at the selected location.

```{r plot gen loc yr one}
do.call(rbind, gen.loc.yr) %>%
  mutate(VAR=factor(VAR)) %>%
  filter(TRIAL==1) %>%
  ggplot() +
    geom_line(aes(SIM_YEAR, VALUE, group=TRIAL), alpha=0.5) +
    geom_line(aes(SIM_YEAR, VALUE), 
              data=mutate(clim.reg.yr, SIM_YEAR=row_number()) %>%
                gather(VAR, VALUE, PRCP:TAVG), 
              color='red') +
    facet_wrap(~VAR, scales='free_y') +
    labs(x="Simulation Year", y="Value", title="Single Generated Dataset with Observed Location Dataset ")
```

The following three figures compare the mean, standard deviation, and skewness of the annual precipitation timeseries between the weather generator simulations and the observed climate dataset at the selected location. The boxplots show the distribution of each statistic across the simulations, the red point shows the mean of the statistic for the simulations, and the blue point shows the statistic value for the observed annual timeseries.

```{r plot gen loc yr box mean, fig.width=4}
do.call(rbind, gen.loc.yr) %>%
  mutate(VAR=factor(VAR)) %>%
  group_by(VAR, TRIAL) %>%
  summarise(MEAN=mean(VALUE)) %>%
  ggplot(aes(x=1, y=MEAN)) +
    geom_boxplot() +
    stat_summary(fun.y=mean, geom="point", color="red", size=3) +
    geom_point(aes(x=1, y=MEAN), 
               data=gather(clim.loc.yr, VAR, VALUE, PRCP:TAVG) %>% 
                      group_by(VAR) %>% 
                      summarise(MEAN=mean(VALUE)),
               color='blue', size=3) +
    facet_wrap(~VAR, scales='free_y') +
    labs(x="", y="Mean", title="Mean of Annual Precipitation for Generated (Red) and Observed (Blue)") +
    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

```{r plot gen loc yr box sd, fig.width=4}
do.call(rbind, gen.loc.yr) %>%
  mutate(VAR=factor(VAR)) %>%
  group_by(VAR, TRIAL) %>%
  summarise(SD=sd(VALUE)) %>%
  ggplot(aes(x=1, y=SD)) +
    geom_boxplot() +
    stat_summary(fun.y=mean, geom="point", color="red", size=3) +
    geom_point(aes(x=1, y=SD), 
               data=gather(clim.loc.yr, VAR, VALUE, PRCP:TAVG) %>% 
                      group_by(VAR) %>% 
                      summarise(SD=sd(VALUE)),
               color='blue', size=3) +
    facet_wrap(~VAR, scales='free_y') +
    labs(x="", y="Standard Deviation", title="StDev of Annual Precipitation for Generated (Red) and Observed (Blue)") +
    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

```{r plot gen loc yr box skew, fig.width=4}
do.call(rbind, gen.loc.yr) %>%
  mutate(VAR=factor(VAR)) %>%
  group_by(VAR, TRIAL) %>%
  summarise(SKEW=skewness(VALUE)) %>%
  ggplot(aes(x=1, y=SKEW)) +
    geom_boxplot() +
    stat_summary(fun.y=mean, geom="point", color="red", size=3) +
    geom_point(aes(x=1, y=SKEW), 
               data=gather(clim.loc.yr, VAR, VALUE, PRCP:TAVG) %>% 
                      group_by(VAR) %>% 
                      summarise(SKEW=skewness(VALUE)),
               color='blue', size=3) +
    facet_wrap(~VAR, scales='free_y') +
    labs(x="", y="Skewness", title="Skewness of Annual Precipitation for Generated (Red) and Observed (Blue)") +
    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

## weatherGen Package

The `weatherGen::gen_mon_arma()` function provides a wrapper around these steps to generate multiple monthly timeseries based on the regional annual precipitation timeseries and the location monthly climate timeseries. This function returns a list object containing the following elements:

- `$models`: list of arma models (only 1 if using ARMA model instead of WARM model)
- `$sim.prcp.yr`: output from `weatherGen::mc_arimas()` containing annual precipitation simulations from the ARMA model
- `$sim.mon`: unnamed list of length `n.iter` with each element containing a data.frame with the generated monthly timeseries for the given location. The columns `YEAR_OBS` and `PRCP_OBS_YEAR` contain the year and corresponding annual precipitation sampled from the monthly timeseries of the location.

```{r wgen mon arma, warning=FALSE}
wgen_mon <- weatherGen::gen_month_arma(x.reg.yr=clim.reg.yr,
                                       x.loc.mon=clim.loc.mon,
                                       n.iter=50, n.year=nrow(clim.reg.yr))
str(wgen_mon, max.level=1)
```

Here is an example of the simulated monthly timeseries for the first trial.

```{r wgen mon example}
head(wgen_mon$sim.mon[[1]])
```

This figure compares the total annual precipitation for the generated monthly timeseries and the annual arma model simulation. Note that the differences between the two reflect the different annual precipitation in the sampled year for the monthly time series and the simulated year from the ARMA model.

```{r plot wgen year compare}
wgen_mon$sim.mon[[1]] %>%
  group_by(YEAR_SIM) %>%
  summarise(PRCP=sum(PRCP),
            YEAR_OBS=mean(YEAR_OBS),
            PRCP_OBS=mean(PRCP_OBS_YEAR)) %>%
  gather(VAR, VALUE, PRCP, PRCP_OBS) %>%
  ggplot(aes(YEAR_SIM, VALUE, color=VAR)) +
  geom_line() +
  labs(x="Year", y="Annual Precipitation (mm)")
```

This figure shows the first simulation trial of the monthly climate timeseries for the selected location.

```{r plot wgen mon}
wgen_mon$sim.mon[[1]] %>%
  select(YEAR_SIM, MONTH, PRCP:TAVG) %>%
  gather(VAR, VALUE, PRCP:TAVG) %>%
  mutate(DATE=ymd(paste(YEAR_SIM, MONTH, 1, sep='-'))) %>%
  ggplot(aes(DATE, VALUE)) +
  geom_line() +
  facet_wrap(~VAR, scales='free_y', ncol=1) +
  labs(x="Month/Year", y="")
```

# Climate Change Trends

After generating the monthly climate timeseries, we can impose linear trends on the timeseries to perform a climate stress test.

## Single Timeseries

The linear trends are incorporated using an additive temperature factor and a multiplicative precipitation factor. For this example, we'll consider an 150% increase in precipitation over the simulation period, and a 2 degC increase in temperature. Note that the same temperature factor is applied to all three temperature variables (TMIN, TMAX, TAVG). We'll impose these trends on only one of the simulated monthly timeseries (chosen at random).

```{r compute trends}
trend.factors <- list(PRCP=1.5, # multiplicative trend factor
                      TEMP=2)   # additive trend factor

# select one trial at random
sim.gen <- wgen_mon$sim.mon[[sample(length(wgen_mon$sim.mon), 1)]]

# create linear trend lines
trend.lines <- list(PRCP=seq(1, trend.factors[['PRCP']], length.out=nrow(sim.gen)),
                    TEMP=seq(0, trend.factors[['TEMP']], length.out=nrow(sim.gen)))

sim.trend <- mutate(sim.gen,
                    PRCP.TREND=PRCP * trend.lines[['PRCP']],
                    TMAX.TREND=TMAX + trend.lines[['TEMP']],
                    TMIN.TREND=TMIN + trend.lines[['TEMP']],
                    TAVG.TREND=TAVG + trend.lines[['TEMP']])
```

This figure compares the original generate timeseries to the adjusted timeseries for the given trend factors.

```{r clim gen trend, fig.height=8}
sim.trend %>%
  mutate(PRCP.GEN=PRCP, TMAX.GEN=TMAX, TMIN.GEN=TMIN, TAVG.GEN=TAVG) %>%
  gather(VAR.GROUP, VALUE, PRCP.GEN:TAVG.GEN, PRCP.TREND:TAVG.TREND) %>%
  separate(VAR.GROUP, c("VAR", "GROUP")) %>%
  mutate(DATE=ymd(paste(YEAR_SIM, MONTH, 1, sep='-'))) %>%
  ggplot(aes(DATE, VALUE, color=GROUP)) +
  geom_line() +
  facet_wrap(~VAR, ncol=1, scales='free_y') +
  labs(x="Month/Year", y="") +
  scale_color_manual('', values=c(GEN='grey20', TREND='chartreuse3'),
                     labels=c(GEN='w/o trend', TREND='w/ trend'))
```

## Multiple Trends for Stress Test

The `weatherGen::trends_mon()` function applies a range of trend factors to the set of  monthly timeseries generated by `weatherGen::gen_month_arma()`. For each pair of temperature and precipitation factors, the function will create a single monthly timeseries by sampling a random iteration from the `gen_month_arma()` output.

```{r}
trend.factors <- list(PRCP=seq(0.75, 1.25, by=0.05),
                      TEMP=seq(0, 5, by=0.5))

sim.trends <- trends_mon(sim.mon.list=wgen_mon$sim.mon, 
                         temp.factors=trend.factors[['TEMP']],
                         prcp.factors=trend.factors[['PRCP']])
str(sim.trends[[1]])
```

Now we can merge these trends and add columns containing the trend factors for temperature (`TEMP_FACTOR`) and precipitaiton (`PRCP_FACTOR`).

```{r trend df}
df.trends <- lapply(sim.trends, function(x) {
  df <- x$df
  df$PRCP_FACTOR <- x$trend.factors[['PRCP']]
  df$TEMP_FACTOR <- x$trend.factors[['TEMP']]
  return(df)
}) %>%
  do.call(rbind, .)
summary(df.trends)
```

This figure shows the monthly precipitation timeseries colored by trend magnitude. This figure only shows precipitation timeseries without any temperature trend. You can see some increase with the 'redder' lines that have higher trend factors.

```{r plot prcp trends}
df.trends %>%
  filter(TEMP_FACTOR==0) %>%
  group_by(PRCP_FACTOR) %>%
  mutate(TIMESTEP=row_number()) %>%
  ggplot(aes(TIMESTEP, PRCP.TREND, color=PRCP_FACTOR)) +
  geom_line(alpha=0.2) +
  scale_color_gradient('Precip Factor', low='green', high='red') +
  labs(x="Simulation Month", y="Monthly Precipitation (mm)")
```

This figure shows the generated precipitation timeseries over the complete range of precipitation trend factors including only timeseries with no temperature trend (`TEMP_FACTOR==0`). This figure clearly shows the increasing trends with higher precipitation trend factors.

```{r plot tavg trends}
df.trends %>%
  filter(TEMP_FACTOR==0) %>%
  group_by(TEMP_FACTOR, PRCP_FACTOR, YEAR_SIM) %>%
  summarise(PRCP=sum(PRCP.TREND)) %>%
  ggplot(aes(YEAR_SIM, PRCP, color=PRCP_FACTOR, group=PRCP_FACTOR)) +
  geom_line(alpha=0.7) +
  scale_color_gradient('Precip Factor', low='green', high='red') +
  labs(x="Simulation Year", y="Annual Precipitation (mm)")
```

This figure shows the generated average temperature timeseries over the range of trend factors including only timeseries with no precipitation trend (`PRCP_FACTOR==0`). This figure clearly shows the increasing trends with higher temperature trend factors.

```{r plot temp trends}
df.trends %>% 
  filter(PRCP_FACTOR==1) %>%
  mutate(DATE=ymd(paste(2001, MONTH, 1, sep='-'))) %>%
  group_by(TEMP_FACTOR, PRCP_FACTOR, YEAR_SIM) %>%
  summarise(TAVG=sum(TAVG.TREND*days_in_month(DATE))/sum(days_in_month(DATE))) %>%
  ggplot(aes(YEAR_SIM, TAVG, color=TEMP_FACTOR, group=TEMP_FACTOR)) +
  geom_line(alpha=0.7) +
  scale_color_gradient('Precip Factor', low='green', high='red') +
  labs(x="Simulation Year", y="Annual Mean Temperature (degC)")
```

# Session Info

```{r session}
sessionInfo()
```


