---
title: "Daily Weather Generator for a Single Site"
author: "Jeffrey D Walker, PhD"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

This document shows how to use the daily weather generator for a single site.

```{r libraries, echo=FALSE, message=FALSE}
library(dplyr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(moments)
library(weathergen)
theme_set(theme_bw())

start_month <- 10
```

# Load Historical Data

First, load the historical daily climate data for Boston, MA from the `climate_cities` dataset included in `weathergen`.

```{r load-obs}
data(climate_cities)
obs_day <- climate_cities[['boston']]

obs_day <- obs_day[weathergen::complete_water_years(obs_day$DATE),]
obs_day['WYEAR'] <- wyear(obs_day[['DATE']], start_month=start_month)
obs_day['MONTH'] <- month(obs_day[['DATE']])
obs_day['TEMP'] <- (obs_day['TMIN'] + obs_day['TMAX'])/2
obs_day <- obs_day[, c('WYEAR', 'MONTH', 'DATE', 'PRCP', 'TEMP', 'TMIN', 'TMAX', 'WIND')]

obs_wyr <- group_by(obs_day, WYEAR) %>%
  summarise(PRCP=sum(PRCP),
            TEMP=mean(TEMP),
            TMIN=mean(TMIN),
            TMAX=mean(TMAX),
            WIND=mean(WIND))

obs <- list(day=obs_day, wyr=obs_wyr)

summary(obs[['day']])
```

This figure shows daily timeseries of each climate variable for Boston, MA.

```{r plot-obs-da, fig.height=8, fig.width=6, echo=FALSE}
gather(obs[['day']], VAR, VALUE, PRCP:WIND) %>%
  ggplot(aes(DATE, VALUE)) +
  geom_line() +
  facet_wrap(~VAR, scales='free_y', ncol=1) +
  labs(x='', y='', title='Historical Daily Weather Data for Boston, MA')
```

This figure shows the annual sum of precipitation, and means of minimum temperature, maximum temperature, and wind speed by water year. 

```{r plot-obs-wyr, fig.height=8, fig.width=6, echo=FALSE}
gather(obs[['wyr']], VAR, VALUE, PRCP:WIND) %>%
  ggplot(aes(WYEAR, VALUE)) +
    geom_line() +
    facet_wrap(~VAR, scales='free_y', ncol=1) +
    labs(x='', y='', title='Historical Annual Weather Data for Boston, MA')
```

# Historical Simulation

In this section, a daily simulation is performed without any adjustment to the statistics or behavior of the weather variables.

## Simulation Parameters

```{r}
start_water_year <- 2000 # starting water year for simulation
n_year <- 40             # number of simulation years
n_knn_annual <- 100      # population size for annual KNN sampling (default)
dry_spell_changes <- 1   # change factor for dry spell (0.8 - 1.2?)
wet_spell_changes <- 1   # change factor for wet spell (0.8 - 1.2?)
prcp_mean_changes <- 1   # change factor for mean precipitation (0.8 - 1.2?)
prcp_cv_changes <- 1     # change factor for precipitation coef. of variation (0.8 - 1.2?)
temp_changes <- 0        # additive factor for temperature variables (-10 - 10)
```

## Annual Precipitation Simulation

A sequence of annual precipitation values are simulated using the `wgen_annual_arima()` function. This function first fits an ARIMA model to the historical annual precipitation. The ARIMA model is then used to generate a simulation of annual precipitation for length `n_year`. The `wgen_annual_arima()` function returns a list of length 3 containing the ARIMA model object (`model`), the historical annual precipitation (`observed`), and the simulated annual precipitation (`values`). 

```{r sim-annual}
sim_annual_prcp <- sim_annual_arima(x = obs[['wyr']][['PRCP']], start_water_year = start_water_year, n_year = n_year)
str(sim_annual_prcp, max.level=1)
```

```{r plot-sim-annual}
data.frame(WYEAR=time(sim_annual_prcp[['sim']]),
           PRCP=zoo::coredata(sim_annual_prcp[['sim']])) %>%
  ggplot(aes(WYEAR, PRCP)) +
  geom_line() +
  geom_hline(yint=mean(obs[['wyr']][['PRCP']]), color='red') +
  geom_text(aes(label=TEXT), data=data.frame(WYEAR=2000, PRCP=mean(obs[['wyr']][['PRCP']]), TEXT="Historical Mean"),
            hjust=0, vjust=-1, color='red') +
  labs(x="Simulation Water Year", y="Annual Precip (mm/yr)")
```

Monte Carlo simulation to compare mean/sd/skew of simulated vs historical (blue point) annual precipitation timeseries.

```{r plot-annual-mc}
# batch simulation to compare mean/sd/skew
z <- lapply(1:50, function(i) { 
    sim_annual_arima(x = obs[['wyr']][['PRCP']], start_water_year = start_water_year, n_year = n_year)$sim
  }) 
names(z) <- paste0(seq(1, length(z)))

do.call(merge, z) %>% 
  apply(MARGIN=2, FUN=function(x) { c(mean=mean(x), sd=sd(x), skew=skewness(x))}) %>% 
  t %>% 
  as.data.frame %>% 
  mutate(trial=row_number()) %>% 
  gather(stat, value, mean:skew) %>% 
  ggplot(aes(x=stat, value)) + 
  geom_boxplot() + 
  geom_point(aes(x=1, y=value), 
             data=data.frame(stat=c('mean', 'sd', 'skew'), 
                             value=c(mean(obs[['wyr']][['PRCP']]), sd(obs[['wyr']][['PRCP']]), skewness(obs[['wyr']][['PRCP']]))), 
             size=4, color='deepskyblue') +
  labs(x='', y='') +
  facet_wrap(~stat, scales='free')
```

## Daily Precipitation Simulation

The annual precipitation generated in the previous step is then used to drive a daily simulation. For each year, a k-nearest neighbors algorithm is used to sample `n_knn_annual` years from the historical record with replacement. A Markov Chain model is then fitted to the sampled historical years and used to simulate a sequence of states for each simulation year. Finally, the simulated states are used to sample the historical daily values using a daily KNN algorithm.

```{r sim-day}
sim_da <- sim_daily_from_annual(prcp_yr = sim_annual_prcp[['sim']],
                                obs_da = zoo(x = obs[['day']][, c('PRCP', 'TEMP', 'TMIN', 'TMAX', 'WIND')], 
                                             order.by = obs[['day']][['DATE']]),
                                obs_prcp_yr = zoo(x = obs[['wyr']][['PRCP']], order.by = obs[['wyr']][['WYEAR']]),
                                start_month = start_month,
                                n_knn_annual = n_knn_annual)
```

```{r plot-sim-day}
select(sim_da, DATE, PRCP, TEMP, TMIN, TMAX) %>%
  gather(VAR, VALUE, PRCP:TMAX) %>%
  ggplot(aes(DATE, VALUE)) +
  geom_line() +
  facet_wrap(~VAR, scales='free_y')
```

Adjust daily simulated precipitation to match annual simulated precipitation.

```{r adj-sim-day}
sim_da_unadj <- select(sim_da, DATE, PRCP)
annual_sum_adjustment <- adjust_daily_to_annual(values_day=sim_da[['PRCP']], 
                                                years_day=wyear(sim_da[['DATE']], start_month = start_month), 
                                                values_yr=coredata(sim_annual_prcp[['sim']]), 
                                                years_yr=time(sim_annual_prcp[['sim']]), 
                                                min_ratio=0.9, max_ratio=1.1)
sim_da[, 'PRCP'] <- annual_sum_adjustment$adjusted
```

Compare the annual precip sum

```{r plot-sim-day-yr-adj}
mutate(sim_da_unadj, WYEAR=wyear(DATE, start_month=start_month)) %>%
  group_by(WYEAR) %>%
  summarise(PRCP=sum(PRCP)) %>%
  mutate(SOURCE="Orig Day") %>%
  rbind(data.frame(WYEAR=time(sim_annual_prcp[['sim']]),
                   PRCP=coredata(sim_annual_prcp[['sim']]),
                   SOURCE="Orig Wyr")) %>%
  rbind(mutate(sim_da, WYEAR=wyear(DATE, start_month=start_month)) %>%
          group_by(WYEAR) %>%
          summarise(PRCP=sum(PRCP)) %>%
          mutate(SOURCE="Adj Day")) %>%
  ggplot(aes(WYEAR, PRCP, color=SOURCE)) +
  geom_line()
```

```{r adj-sim-day-changes}
# TODO: compute mean_change from transition matrices
gamma_adjustment <- adjust_daily_gamma(sim_da[['PRCP']], sim_da[['MONTH']], mean_change=prcp_mean_changes, cv_change=prcp_cv_changes)
sim_da[, 'PRCP'] <- gamma_adjustment$adjusted

tmin_adjustment <- adjust_daily_additive(x = sim_da[['TMIN']], months = sim_da[['MONTH']], changes=temp_changes)
tmax_adjustment <- adjust_daily_additive(x = sim_da[['TMAX']], months = sim_da[['MONTH']], changes=temp_changes)
temp_adjustment <- adjust_daily_additive(x = sim_da[['TEMP']], months = sim_da[['MONTH']], changes=temp_changes)
sim_da[, 'TMIN'] <- tmin_adjustment$adjusted
sim_da[, 'TMAX'] <- tmax_adjustment$adjusted
sim_da[, 'TEMP'] <- temp_adjustment$adjusted
```


```{r, eval=FALSE}
rbind(select(sim_da, DATE, PRCP) %>% mutate(SOURCE="Adj"),
      select(sim_da_unadj, DATE, PRCP) %>% mutate(SOURCE="Orig")) %>%
  mutate(WYEAR=wyear(DATE, start_month=start_month)) %>%
  select(WYEAR, DATE, SOURCE, PRCP) %>%
  spread(SOURCE, PRCP) %>%
  mutate(DIFF=Adj-Orig,
         RATIO=ifelse(Orig==0, NA, Adj/Orig)) %>%
  ggplot(aes(DATE, RATIO)) +
  geom_point() +
  geom_hline(yint=1)
```


